---
title: "Example 0"
author: "Richard J Beck"
date: "2023-10-10"
output:
  md_document:
    variant: markdown_github
---

**This example shows how to visualize karyotype frequencies from the input data and overlay ALFA-K fit results on the plots.**

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/projects/008_birthrateLandscape/ALFA-K/")
```

```{r,warning=FALSE,message=FALSE}
library(ggplot2)
source("utils/visualisation_functions.R")
```

melt_for_plotting() plots karyotype frequencies over time, with option to overlay alfa-k fitted frequency estimates. See example_1 for more info on generating suitable input data for this function.

**Arguments**

**data_obj:** karyotype frequency data formatted as output from proc_sim() function

**nclones:** number of karyotypes to plot (sorted by descending frequency in input data)

**fit_obj:** (optional) fit resulting from alfak() function

**Value**

melt_for_plotting returns a list that is convenient for input into ggplot

**\$data:** data frame containing karyotype frequencies from input data

**\$fit:** either NULL or a dataframe containing karyotype frequency estimates from ALFA-K

example:

```{r}

data_obj <- readRDS("example_data/SA535.Rds")
fit_obj <- readRDS("example_data/SA535_landscape.Rds")

x1 <- melt_for_plotting(data_obj,nclones=5)
x2 <- melt_for_plotting(data_obj,nclones=5,fit_obj)


p <- ggplot(x1$data,aes(x=time,y=frequency,color=karyotype))+
  geom_point(size=2)+
  geom_line()+
  theme_classic(base_size=12)
p

p <- ggplot(x2$data,aes(x=time,y=frequency,color=karyotype))+
  geom_point(size=2)+
  geom_line(data=x2$fit)+
  theme_classic(base_size=12)
p


```
