---
title: "README"
author: "Richard J Beck"
date: "`r Sys.Date()`"
output: html_document
---

Validating ALFA-K against output of ABM simulations  

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/projects/008_birthrateLandscape/ALFA-K/")
```

```{r}
library(ggplot2)
```

INCLUDE DETAILS ABOUT HOW ABM OUTPUT WAS GENERATED

First provide some background info about the sims which might help in interpreting results:

```{r,eval=F}
source("figures/alfak_ABM_tests/extract_sim_summary_info.R") ## ran on workstation
```

First check the fitness changes that occured in the ABM simulations. 
```{r}
x <- readRDS("figures/alfak_ABM_tests/sim_summary_info.Rds")
ids <- data.frame(do.call(rbind,lapply(x$cond_id, function(i) unlist(strsplit(i,split="_"))[c(4,6,8)])))
colnames(ids) <- c("wavelength","misrate","rep")
x <- cbind(x,ids)
x <- x[x$misrate=="0.00005",]
p <- ggplot(x,aes(x=time,y=fitness,group=interaction(rep_id,cond_id)))+
  facet_grid(cols=vars(wavelength))+
  geom_line(alpha=0.1)
p

```


```{r}

x <- readRDS("figures/alfak_ABM_tests/fit_summary_info.Rds")
ids <- data.frame(do.call(rbind,lapply(x$id1, function(i) unlist(strsplit(i,split="_"))[c(4,6,8)])))
colnames(ids) <- c("wavelength","misrate","rep")
x <- cbind(x,ids)
x <- x[x$misrate=="0.00005",]

ids <- data.frame(do.call(rbind,lapply(x$id2, function(i) unlist(strsplit(i,split="_"))[c(2,4)])))
colnames(ids) <- c("min_obs","ntp")
x <- cbind(x,ids)                   
x$ntp <- paste(x$ntp,"samples")
x$min_obs <- paste0("min.samples=",stringr::str_pad(x$min_obs,width = 2))
x <- x[,!colnames(x)%in%c("id1","id2")]

p <- ggplot(x,aes(x=wavelength,y=cor,color=id))+
  facet_grid(rows=vars(min_obs),cols=vars(ntp))+
  geom_jitter(height=0,width=0.2,alpha=0.5)
p
```

```{r}

x <- readRDS("figures/alfak_ABM_tests/fit_summaries.Rds")
x <- x[!sapply(x,is.null)]
x <- x[sapply(x,ncol)==6]
x <- do.call(rbind,x)
ids <- data.frame(do.call(rbind,lapply(x$cond_id, function(i) unlist(strsplit(i,split="_"))[c(4,6,8)])))
colnames(ids) <- c("wavelength","misrate","rep")
x <- cbind(x,ids)
x <- x[x$misrate=="0.00005",]
x$ntp <- paste(x$ntp,"samples")
x$min_obs <- paste0("min.samples=",stringr::str_pad(x$min_obs,width = 2))

p <- ggplot(x,aes(x=wavelength,y=r2,color=id))+
  facet_grid(rows=vars(min_obs),cols=vars(ntp))+
  geom_jitter(height=0,width=0.2,alpha=0.5)
p

```

```{r}

s <- readRDS("figures/alfak_ABM_tests/sim_summary_info.Rds")
#s <- s[grepl("N_22_w_0p8",s$cond_id),]
s <- s[s$time>2000&s$time<2200,]
s <- split(s,f=interaction(s$cond_id,s$rep_id))

ss <- do.call(rbind,lapply(s,function(si){
  if(nrow(si)<2) return(data.frame(cond_id=si$cond_id[1],rep_id=si$rep_id[1],delta_f=NaN))
  delta_f <- lm(fitness~time,si)$coefficients[2]
  data.frame(cond_id=si$cond_id[1],rep_id=si$rep_id[1],delta_f)
}))

ss <- ss[grepl("0.00005",ss$cond_id),]
rownames(ss) <- interaction(ss$rep_id,ss$cond_id)
```

```{r}

fs <- readRDS("figures/alfak_ABM_tests/fit_summary_info.Rds")
fs <- fs[fs$id=="n2",]
fs$id2 <- gsub(".Rds","",fs$id2)
rownames(fs)<- interaction(fs$id1,fs$id2)
```

```{r}

xv <- readRDS("figures/alfak_ABM_tests/loo_summaries.Rds")
#ids <- data.frame(do.call(rbind,lapply(xv$rep, function(i) unlist(strsplit(i,split="_"))[c(4,6,8)])))
#colnames(ids) <- c("wavelength","misrate","rep")
#xv <- cbind(xv,ids)
xv <- split(xv,f=interaction(xv$ntp,xv$minobs,xv$rep))
xv <- do.call(rbind,lapply(xv,function(xi){
  
  dfi <- xi[1,c("rep","ntp","minobs")]
  
dfi$cor <- tryCatch(R2(xi$pred[!is.na(xi$pred)],xi$f_est[!is.na(xi$pred)]),
                    error=function(e) return(NaN))
return(dfi)
}))
xv$cor[xv$cor<(-1)]<- -1
xv$cor[!is.finite(xv$cor)] <- -1
xv$condition <- paste0("minobs_",xv$minobs,"_ntp_",xv$ntp,"_00000")
rownames(xv) <- interaction(xv$condition,xv$rep)

```

```{r}
wavelength <- "1p6"
predictor <- "XV score"
#predictor <- "charted size"
#predictor <- "fitness change"

x <- readRDS("figures/alfak_ABM_tests/fitted_abm_run_angles.Rds")
ids <- data.frame(do.call(rbind,lapply(x$id, function(i) unlist(strsplit(i,split="_"))[c(4,6,8)])))
colnames(ids) <- c("wavelength","misrate","rep2")
x <- cbind(x,ids)
ids <- data.frame(do.call(rbind,lapply(x$condition, function(i) unlist(strsplit(i,split="_"))[c(2,4)])))
colnames(ids) <- c("min_obs","ntp")
x <- cbind(x,ids)
x$ntp <- paste(x$ntp,"samples")
x$min_obs <- paste0("min.samples=",stringr::str_pad(x$min_obs,width = 2))

y <- x[x$wavelength==wavelength,]

y$predictor <- NaN

if(predictor=="charted size") y$predictor <- fs$n[interaction(y$id,y$condition)]
if(predictor=="XV score") y$predictor <- xv$cor[interaction(y$condition,y$id)]
if(predictor=="fitness change") y$predictor <- ss$delta_f[interaction(y$rep,y$id)]


y <- split(y,f=interaction(y$min_obs,y$ntp))
y <- do.call(rbind,lapply(y,function(yi){
  pp <- rep("low",nrow(yi))
  
  if(predictor=="XV score"){
    pp[yi$predictor>0.45] <- "high"
  }else{
    pp[yi$predictor>quantile(yi$predictor,p=0.5)] <- "high"
  }
  yi$predictor <- pp
  return(yi)
}))


plookup <- seq(2100,2900,100)
names(plookup) <- paste0("t",seq(2100,2900,100))
y <- reshape2::melt(y,measure.vars=paste0("t",seq(2100,2900,100)))
z <- aggregate(list(mn=y$value),by=list(variable=y$variable,
                                        min_obs=y$min_obs,
                                        ntp=y$ntp,
                                        predictor=y$predictor),
               mean)

z2 <- aggregate(list(sd=y$value),by=list(variable=y$variable,
                                        min_obs=y$min_obs,
                                        ntp=y$ntp,
                                        predictor=y$predictor),
               sd)

z$sd <- z2$sd

p <- ggplot(z,aes(x=(plookup[variable]-2000)/10,y=mn,color=predictor))+
  facet_grid(rows=vars(min_obs),cols=vars(ntp))+
  geom_point()+
  scale_color_discrete(predictor)+
  geom_errorbar(aes(ymin=mn-sd,ymax=mn+sd))+
  scale_x_continuous("time into future (days)")+
  scale_y_continuous("angle metric")+
  ggtitle(paste("wavelength:",wavelength))
p

```

```{r}
wavelength <- "0p8"
predictor <- "XV score"
#predictor <- "charted size"
#predictor <- "fitness change"

x <- readRDS("figures/alfak_ABM_tests/fitted_abm_run_angles.Rds")
ids <- data.frame(do.call(rbind,lapply(x$id, function(i) unlist(strsplit(i,split="_"))[c(4,6,8)])))
colnames(ids) <- c("wavelength","misrate","rep2")
x <- cbind(x,ids)
ids <- data.frame(do.call(rbind,lapply(x$condition, function(i) unlist(strsplit(i,split="_"))[c(2,4)])))
colnames(ids) <- c("min_obs","ntp")
x <- cbind(x,ids)
x$ntp <- paste(x$ntp,"samples")
x$min_obs <- paste0("min.samples=",stringr::str_pad(x$min_obs,width = 2))

y <- x[x$wavelength==wavelength,]
y <- x
y$predictor <- NaN

if(predictor=="charted size") y$predictor <- fs$n[interaction(y$id,y$condition)]
if(predictor=="XV score") y$predictor <- xv$cor[interaction(y$condition,y$id)]
if(predictor=="fitness change") y$predictor <- ss$delta_f[interaction(y$rep,y$id)]


y <- split(y,f=interaction(y$min_obs,y$ntp))
y <- do.call(rbind,lapply(y,function(yi){
  pp <- rep("low",nrow(yi))
  
  if(predictor=="XV score"){
    pp[yi$predictor>0.45] <- "high"
  }else{
    pp[yi$predictor>quantile(yi$predictor,p=0.5,na.rm=T)] <- "high"
  }
  yi$predictor <- pp
  return(yi)
}))


plookup <- seq(2100,2900,100)
names(plookup) <- paste0("t",seq(2100,2900,100))
y <- reshape2::melt(y,measure.vars=paste0("t",seq(2100,2900,100)))
z <- aggregate(list(mn=y$value),by=list(variable=y$variable,
                                        min_obs=y$min_obs,
                                        ntp=y$ntp,
                                        predictor=y$predictor),
               mean)

z2 <- aggregate(list(sd=y$value),by=list(variable=y$variable,
                                        min_obs=y$min_obs,
                                        ntp=y$ntp,
                                        predictor=y$predictor),
               sd)

z$sd <- z2$sd

p <- ggplot(z,aes(x=(plookup[variable]-2000)/10,y=mn,color=predictor))+
  facet_grid(rows=vars(min_obs),cols=vars(ntp))+
  geom_point()+
  scale_color_discrete(predictor)+
  geom_errorbar(aes(ymin=mn-sd,ymax=mn+sd))+
  scale_x_continuous("time into future (days)")+
  scale_y_continuous("angle metric")
  #ggtitle(paste("wavelength:",wavelength))
p

```

```{r}

x <- readRDS("figures/alfak_ABM_tests/fitted_abm_run_angles.Rds")
ids <- data.frame(do.call(rbind,lapply(x$id, function(i) unlist(strsplit(i,split="_"))[c(4,6,8)])))
colnames(ids) <- c("wavelength","misrate","rep2")
x <- cbind(x,ids)
ids <- data.frame(do.call(rbind,lapply(x$condition, function(i) unlist(strsplit(i,split="_"))[c(2,4)])))
colnames(ids) <- c("min_obs","ntp")
x <- cbind(x,ids)
x$ntp <- paste(x$ntp,"samples")
x$min_obs <- paste0("min.samples=",stringr::str_pad(x$min_obs,width = 2))

z <- reshape2::melt(x,measure.vars=paste0("p",1:5))
z <- z[z$min_obs == "min.samples=10", ]



p3 <- ggplot(x,aes(x=wavelength,y=p3))+
  facet_grid(rows=vars(min_obs),cols=vars(ntp))+
  geom_violin()+
  geom_jitter(height=0,width=0.3,alpha=0.2)+
  geom_hline(yintercept=90,color="red")
p3

pt <- ggplot(z,aes(x=variable,y=value))+
  facet_grid(rows=vars(wavelength),cols=vars(ntp))+
  geom_violin()+
  #geom_jitter(height=0,width=0.3,alpha=0.2)+
  geom_hline(yintercept=90,color="red")+
  scale_x_discrete("passage number")
pt


```


```{r}
x <- readRDS("figures/alfak_ABM_tests/loo_summaries.Rds")
ids <- data.frame(do.call(rbind,lapply(x$rep, function(i) unlist(strsplit(i,split="_"))[c(4,6,8)])))
colnames(ids) <- c("wavelength","misrate","rep")
x <- cbind(x,ids)

x <- split(x,f=interaction(x$ntp,x$minobs,x$rep))

df <- do.call(rbind,lapply(x,function(xi){
  
  dfi <- xi[,c("ntp","minobs","wavelength","misrate")]
  
dfi$cor <- tryCatch(R2(xi$pred[!is.na(xi$pred)],xi$f_est[!is.na(xi$pred)]),
                    error=function(e) return(NaN))
return(dfi)
}))
df$cor[df$cor<(-1)]<- -1
p <- ggplot(df,aes(x=wavelength,y=cor))+
  facet_grid(rows=vars(minobs),cols=vars(ntp))+
  geom_violin()+
  geom_jitter(height=0,width=0.3,alpha=0.1)+
  scale_y_continuous(expression(R^2))
p

```


I think reason that the model doesn't predict very well for just 2 samples is that it becomes really hard to constrain the fit. Note that in the following, the fits and Pearson correlation values are pretty good even when the relative fitness estimates are way off. 
```{r}
set.seed(42)
run_model <- function(x0,g,tt){
  u <- exp(g%*%t(tt))
  for(i in 1:nrow(u)) u[i,] <- u[i,]*exp(x0[i])
  x <- data.frame(t(apply(u,2,function(ui) ui/sum(ui))))
  x$time <- tt
  x <- reshape2::melt(x,id.vars="time")
  x <- x[order(x$time,x$variable),] 
  return(x)
}

optimr <- function(pars,y){
  npars <- length(pars)/2
  x0 <- pars[1:npars]
  g <- pars[(npars+1):length(pars)]
  tt <- unique(y$time)
  x <- run_model(x0,g,tt)
  y <- y[order(y$time,y$variable),]
  
  sqrt(mean((x$value-y$value)^2))
  
}

x0 <- c(-20,-10,-5)
g <- c(.8,0.65,.5)
tt <- 0:100

u <- exp(g%*%t(tt))
for(i in 1:nrow(u)) u[i,] <- u[i,]*exp(x0[i])

x <- run_model(x0,g,tt)
y <- x[x$time%in%c(80,20),]


p1 <- ggplot(x,aes(x=time,y=value,color=variable))+
  geom_line()+
  geom_point(data=y,size=3)+
  ggtitle("input data + perfect samples")+
  scale_color_discrete("karyotype")
p1

z <- pbapply::pblapply(1:50, function(i){
  pars <- c(runif(length(x0),-20,0),runif(length(g),0,1))
  opt <- optim(pars,fn = optimr,y=y)
  pars <- opt$par
  npars <- length(pars)/2
  x0 <- pars[1:npars]
  g <- pars[(npars+1):length(pars)]
  zi <- run_model(x0,g,tt)
  zi$id <- i
  zi$err <- opt$value
  
  zopt <- data.frame(g=tail(opt$par,npars),id=c("X1","X2","X3"),rep=i,err=opt$value)
  zopt$g <- zopt$g-min(zopt$g)
  return(list(z=zi,zopt=zopt))
})
zopt <- do.call(rbind,lapply(z,function(zi) zi$zopt))
z <- do.call(rbind,lapply(z,function(zi) zi$z))
z[,] <- lapply(z, function(x) type.convert(as.character(x), as.is = TRUE))
#zplt <- z[z$]
p2 <- ggplot(z[z$err<0.05,],aes(x=time,y=value,color=variable))+
  geom_line(aes(group=interaction(variable,id)))+
  geom_point(data=y,size=3)+
  ggtitle("fits to samples")+
  scale_color_discrete("karyotype")
p2

p3 <- ggplot(zopt[zopt$err<0.05,],aes(x=id,y=g,group=rep))+
  geom_point()+
  geom_line()+
  scale_y_continuous("estimated growth rate")+
  scale_x_discrete("")
p3

tmp <- split(zopt[zopt$err<0.05,],f=zopt[zopt$err<0.05,]$rep)
df <- do.call(rbind,lapply(tmp, function(ti) {
  data.frame(r2=cor(ti$g,g)*abs(cor(ti$g,g)),R2=R2(ti$g,g-min(g)),rep=ti$rep[1])
  }))
df$R2[df$R2<(-1)]<-(-1)

df <- reshape2::melt(df,id.vars=c("rep"))

p4 <- ggplot(df,aes(x=value,fill=variable))+
  geom_histogram(binwidth=0.1,position="dodge")
p4

```