---
title: "README"
author: "Richard J Beck"
date: "`r Sys.Date()`"
output: html_document
---

Validating ALFA-K against output of ABM simulations  

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/projects/008_birthrateLandscape/ALFA-K/")
```

```{r}
library(ggplot2)
source("utils/comparison_functions.R")
```

INCLUDE DETAILS ABOUT HOW ABM OUTPUT WAS GENERATED

First provide some background info about the sims which might help in interpreting results:

```{r,eval=F}
source("figures/alfak_ABM_tests/extract_sim_summary_info.R") ## ran on workstation
```

First check the fitness changes that occured in the ABM simulations. 
```{r}
x <- readRDS("figures/alfak_ABM_tests/sim_summary_info.Rds")
ids <- data.frame(do.call(rbind,lapply(x$cond_id, function(i) unlist(strsplit(i,split="_"))[c(4,6,8)])))
colnames(ids) <- c("wavelength","misrate","rep")
x <- cbind(x,ids)
x <- x[x$misrate=="0.00005",]
p <- ggplot(x,aes(x=time/10,y=fitness,group=interaction(rep_id,cond_id)))+
  facet_grid(cols=vars(paste0("lambda==",gsub("p",".",wavelength))),labeller = label_parsed)+
  geom_line(alpha=0.1)+
  scale_x_continuous("days")+
  scale_y_continuous("mean fitness")+
  geom_vline(xintercept=200,color="red",linetype=2)
p

ggsave("figures/alfak_ABM_tests/figs/simulation_fitness.png",width=7,height=2,units="in")

```

```{r}
x <- readRDS("figures/alfak_ABM_tests/loo_summaries.Rds")
ids <- data.frame(do.call(rbind,lapply(x$rep, function(i) unlist(strsplit(i,split="_"))[c(4,6,8)])))
colnames(ids) <- c("wavelength","misrate","rep")
x <- cbind(x,ids)

x <- split(x,f=interaction(x$ntp,x$minobs,x$rep))

df <- do.call(rbind,lapply(x,function(xi){
  
  dfi <- xi[,c("ntp","minobs","wavelength","misrate")]
  
dfi$cor <- tryCatch(R2(xi$pred[!is.na(xi$pred)],xi$f_est[!is.na(xi$pred)]),
                    error=function(e) return(NaN))
return(dfi)
}))
df$cor[df$cor<(-1)]<- -1
p <- ggplot(df,aes(x=gsub("p",".",wavelength),y=cor))+
  facet_grid(rows=vars(paste0("N=",stringr::str_pad(minobs,width = 2))),cols=vars(paste0(ntp," sampled timepoints")))+
  geom_violin()+
  #geom_jitter(height=0,width=0.3,alpha=0.1)+
  scale_y_continuous(expression(cross~validation~R^2))+
  scale_x_discrete(expression(GRF~wavelength~(lambda)))
p

ggsave("figures/alfak_ABM_tests/figs/validation_r2.png",width=6,height=4,units="in")

```


```{r}

x <- readRDS("figures/alfak_ABM_tests/fitted_abm_run_angles.Rds")
ids <- data.frame(do.call(rbind,lapply(x$id, function(i) unlist(strsplit(i,split="_"))[c(4,6,8)])))
colnames(ids) <- c("wavelength","misrate","rep2")
x <- cbind(x,ids)
ids <- data.frame(do.call(rbind,lapply(x$condition, function(i) unlist(strsplit(i,split="_"))[c(2,4)])))
colnames(ids) <- c("min_obs","ntp")
x <- cbind(x,ids)
x$ntp <- paste(x$ntp,"sampled timepoints")
x$min_obs <- paste0("N=",stringr::str_pad(x$min_obs,width = 2))

z <- reshape2::melt(x,measure.vars=paste0("t",seq(2100,2900,100)))
#z <- z[z$min_obs == "min.samples=10", ]



p3 <- ggplot(z[z$variable=="t2500",],aes(x=gsub("p",".",wavelength),
                   y=value))+
  facet_grid(rows=vars(min_obs),cols=vars(ntp))+
  geom_violin()+
  geom_hline(yintercept=90,color="red")+
  scale_fill_discrete(expression(lambda))+
  scale_y_continuous("angle metric")+
  scale_x_discrete(expression(GRF~landscape~wavelength~(lambda)))
p3
zf <- z[z$wavelength=="0p8"&z$min_obs=="N=10"&z$ntp=="4 sampled timepoints",]
p3b <- ggplot(zf,aes(x=as.numeric(gsub("t","",variable))/10,y=value,group=variable))+
  geom_violin()+
  scale_x_continuous(breaks=seq(200,300,20))+
  scale_y_continuous("angle metric")
p3b

ggsave("figures/alfak_ABM_tests/figs/validation_angles.png",width=6,height=4,units="in")
```
Agreement between R^2 and angle metric scores.
```{r}
xv <- readRDS("figures/alfak_ABM_tests/loo_summaries.Rds")
xv <- split(xv,f=interaction(xv$ntp,xv$minobs,xv$rep))
xv <- do.call(rbind,lapply(xv,function(xi){
  dfi <- xi[1,c("rep","ntp","minobs")]
dfi$cor <- tryCatch(R2(xi$pred[!is.na(xi$pred)],xi$f_est[!is.na(xi$pred)]),
                   error=function(e) return(NaN))
return(dfi)
}))
xv$cor[xv$cor<(-1)]<- -1
xv$cor[!is.finite(xv$cor)] <- -1
xv$condition <- paste0("minobs_",xv$minobs,"_ntp_",xv$ntp,"_00000")
rownames(xv) <- interaction(xv$condition,xv$rep)

```

```{r}
x <- readRDS("figures/alfak_ABM_tests/fitted_abm_run_angles.Rds")
ids <- data.frame(do.call(rbind,lapply(x$id, function(i) unlist(strsplit(i,split="_"))[c(4,6,8)])))
colnames(ids) <- c("wavelength","misrate","rep2")
x <- cbind(x,ids)
ids <- data.frame(do.call(rbind,lapply(x$condition, function(i) unlist(strsplit(i,split="_"))[c(2,4)])))
colnames(ids) <- c("min_obs","ntp")
x <- cbind(x,ids)
x$ntp <- paste(x$ntp,"samples")
x$min_obs <- paste0("min.samples=",stringr::str_pad(x$min_obs,width = 2))

x$predictor <- xv$cor[interaction(x$condition,x$id)]

x <- split(x,f=interaction(x$min_obs,x$ntp))
x <- do.call(rbind,lapply(x,function(yi){
  pp <- rep("low",nrow(yi))
  pp[yi$predictor>0] <- "high"
  yi$predictor <- pp
  return(yi)
}))


plookup <- seq(2100,2900,100)
names(plookup) <- paste0("t",seq(2100,2900,100))
x <- reshape2::melt(x,measure.vars=paste0("t",seq(2100,2900,100)))


p <- ggplot(x,aes(x=as.character(as.numeric(gsub("t","",variable))/10),y=value,fill=predictor))+
  geom_violin()+
  scale_fill_discrete(expression(R^2))+
  scale_x_discrete("time (days)")+
  scale_y_continuous("angle metric",breaks=seq(0,180,45))
p

ggsave("figures/alfak_ABM_tests/figs/r2_angles.png",width=4,height=2,units="in")
```


Test how missegregation rate affects fitted results:

Input data for this sweep was generated using the following command:
Rscript generate_ABM_sweep.R -n data/misrate -r 100 -w 0.8 -m 0.00001,0.00003,0.0001,0.0003,0.001 
Rscript process_abm_sweep.R -n data/misrate 
Rscript sweep_summary_stats -n data/misrate -o data/misrate_stats.Rds

```{r}
x <- readRDS("data/misrate_stats.Rds")
x$ll_fq <- -x$ll_fq
x$ll_nn <- -x$ll_nn
df <- reshape2::melt(x[,c("m","nfq","r_fq","r_nn","r_d2n")],id.vars=c("m","nfq"))
ids <- c(" frequent","d1","d2")
names(ids)<-c("r_fq","r_nn","r_d2n")
df$variable <- ids[df$variable]
df$value[df$value<(-1)] <- -1
p0 <- ggplot(df,aes(x=factor(m),y=value))+
  facet_grid(rows=vars(variable))+
  geom_violin()+
  geom_jitter(height=0,width=0.2)+
  scale_y_continuous(expression(predicted~fitness~R^2))+
  scale_x_discrete("missegregation rate")
p0

df <- reshape2::melt(x[,c("m","r_fq","nfq","nnn","nd2n")],id.vars=c("m","r_fq"))
ids <- c(" frequent","d1","d2")
names(ids)<-c("nfq","nnn","nd2n")
df$variable <- ids[df$variable]
df$r_fq[df$r_fq<(-1)] <- -1
p1 <- ggplot(df,aes(x=factor(m),y=value,color=r_fq))+
  facet_grid(rows=vars(gsub("d","distance ",variable)),scales="free")+
  geom_violin()+
  geom_jitter(height=0,width=0.2)+
  scale_y_log10("number of karyotypes")+
  scale_x_discrete("missegregation rate")+
  scale_color_viridis_c(expression(R^2))
p1
ggsave("figures/alfak_ABM_tests/figs/misseg_sweep.png",width=5,height=3,units="in")
```

