---
title: "README"
author: "Richard J Beck"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir="~/projects/008_birthrateLandscape/ALFA-K/")
```

2 modes of failure:

1) No evolution
2) No enough samples

We will explore what happens when there are not enough samples first.


```{r}
root.dir <- "data/main/"
sweepID <- "N_22_w_1p6_m_0.00005_rep_08"
dir <- paste0(root.dir,sweepID,"/")

z <- readRDS("tests/data/pearsonR.Rds")
z <- z[z$id1==sweepID,]
z <- z[grepl("_10_",z$id2),]

id <- z$id2[1]
fit_obj <- readRDS(paste0(dir,"sweep_fits/",id,".Rds"))
fobj <- gen_fitness_object(paste0(dir,"/config.txt"), paste0(dir,"/landscape.txt"))
k <- head(rownames(fit_obj$xo),5)
kmat <- do.call(rbind,lapply(k,s2v))
ftru <- apply(kmat,1,getf,fobj=fobj)


q <- lapply(z$id2,function(id){
  ntp <- as.numeric(unlist(strsplit(id,"_"))[4])
  fit_obj <- readRDS(paste0(dir,"sweep_fits/",id,".Rds"))
  data_obj <- proc_sim(paste0(dir,"train/00000/"),seq(0,2000,length.out=ntp))
  q <- melt_for_plotting(data_obj=data_obj,karyotypes = k,fit_obj = fit_obj)
  q$est <- data.frame(f_est=fit_obj$xo[k,]$f_est,f_tru=ftru,ntp=ntp)
  q$data$ntp <- ntp
  q$fit$ntp <- ntp
  q
})

qdata <- do.call(rbind,lapply(q,function(qi) qi$data))
qfit <- do.call(rbind,lapply(q,function(qi) qi$fit))
qest <-  do.call(rbind,lapply(q,function(qi) qi$est))

```

In this figure we're comparing a situation where we sample the same ABM experiment at either 2,4,or 8 timepoints. Key points:
1) Note all of the fits to the available data are good.
2) Clearly with 2 timepoints the fit is underconstrained. 
 
```{r}

p1 <- ggplot(qdata,aes(x=time,y=frequency,color=karyotype))+
  facet_wrap(~ntp)+
  geom_point(size=2)+
  geom_line(data=qfit)+
  theme_classic(base_size=12)
p1

```

As we examine the clone fitness estimates resulting from the fit we notice:
1) for 2 timepoints case there are two clumps of fitness estimates - 1 of clones present at initial timepoint (in this case a single, diploid clone), another of the clones that emerge at later timepoints.
2) Adding more timepoints the fitness estimates correlate better with the true fitness.
3) Note in the 4 timepoints case the bad estimate - results from the underconstrained clone that is emerging at the final timepoint. Going from 4-8 timepoints one can observe how this clone fitness is better constrained.


```{r}

p2 <- ggplot(qest,aes(x=f_est,y=f_tru,color=as.character(ntp)))+
  theme_classic(base_size=12)+
  geom_abline(intercept=0.28)+
  geom_jitter(height=0.002,width=0.002,size=2)
p2
```

```{r}


root.dir <- "data/main/"
sweepID <- "N_22_w_1p6_m_0.00005_rep_08"
dir.create(paste0("figures/worked_examples/data/",sweepID),showWarnings = F)
dir <- paste0(root.dir,sweepID,"/")

z <- readRDS("tests/data/pearsonR.Rds")
z <- z[z$id1==sweepID,]
z <- z[grepl("_10_",z$id2),]

fobj <- gen_fitness_object(paste0(dir,"/config.txt"), paste0(dir,"/landscape.txt"))



q <- do.call(rbind,lapply(z$id2,function(id){
  ntp <- as.numeric(unlist(strsplit(id,"_"))[4])
  
  x <- proc_sim(paste0(dir,"train/00000/"),seq(0,2000,length.out=ntp))
  xo <- alfak(x,min_obs = 10)
  n <- do.call(rbind,lapply(rownames(xo$xo),s2v))
  xo$xo$f_tru <- apply(n,1,getf,fobj=fobj)
  
  n2 <- gen_all_neighbours(rownames(xo$xo))
  ftru2 <- apply(n2,1,getf,fobj=fobj)
  fest2 <- predict(xo$fit,n2)
  
  df2 <- data.frame(f_est=fest2,f_tru=ftru2)
  
  xo$d2n <- df2
  
  saveRDS(xo,paste0("figures/worked_examples/data/",sweepID,"/",id,".Rds"))
}))

dd <- "figures/worked_examples/data/N_22_w_1p6_m_0.00005_rep_08/"
for(d in list.files(dd)){
  x <- readRDS(paste0(dd,d))
  n <- do.call(rbind,lapply(rownames(x$xo),s2v))
  x$xo$f_tru <- apply(n,1,getf,fobj=fobj)
  colnames(x$d2n) <- c("f_tru","f_est")
  saveRDS(x,paste0(dd,d))
}
```



```{r}

dd <- "figures/worked_examples/data/N_22_w_1p6_m_0.00005_rep_08/"
f <- list.files(dd)

y <- do.call(rbind,lapply(f,function(fi){
  x <- readRDS(paste0(dd,fi))
  x$d2n$id <- "d2n"
  y <- rbind(x$xo[,colnames(x$d2n)],x$d2n)
  y$ntp <- unlist(strsplit(fi,split="_"))[4]
  y
}))


p <- ggplot(y,aes(x=f_est,y=f_tru))+
  facet_grid(cols=vars(ntp),rows=vars(id),scales="free")+
  geom_point(aes(color=id))
p

```

```{r}


root.dir <- "data/main/"
sweepID <- "N_22_w_1p6_m_0.00005_rep_08"
dir <- paste0(root.dir,sweepID,"/")

id <- "minobs_10_ntp_2_00000"
outDir <- paste0("figures/worked_examples/data/",id,"/")
dir.create(outDir)
set.seed(42)


fobj <- gen_fitness_object(paste0(dir,"/config.txt"), paste0(dir,"/landscape.txt"))

reps <- 0:9
ntp <- as.numeric(unlist(strsplit(id,"_"))[4])
x <- proc_sim(paste0(dir,"train/00000/"),seq(0,2000,length.out=ntp))
for(rep in reps){
  xo <- alfak(x,min_obs = 10)
  n <- do.call(rbind,lapply(rownames(xo$xo),s2v))
  xo$xo$f_tru <- apply(n,1,getf,fobj=fobj)
  saveRDS(xo,paste0(outDir,"test_",rep,".Rds"))
  print(cor(xo$xo$f_est[xo$xo$id=="fq"],xo$xo$f_tru[xo$xo$id=="fq"]))
  print(cor(xo$xo$f_est[xo$xo$id=="nn"],xo$xo$f_tru[xo$xo$id=="nn"]))
}



```


```{r}

root.dir <- "data/main/"
sweepID <- "N_22_w_1p6_m_0.00005_rep_08"
dir <- paste0(root.dir,sweepID,"/")

id <- "minobs_10_ntp_2_00000"
ntp <- as.numeric(unlist(strsplit(id,"_"))[4])

x <- proc_sim(paste0(dir,"train/00000/"),seq(0,2000,length.out=ntp))

x0 <- readRDS("figures/worked_examples/data/minobs_10_ntp_2_00000/test_1.Rds")
x0 <- x0$xo[x0$xo$id=="fq",]
x1 <- wrap_neighbor_fitness(x,x0,pm0=misseg_rate)
hist(x1$f_est)
```

Seems like something is different about our neighbor fitness estimation that means our old result is not reproducible. Plus side - new result looks better. Guess we better rerun.

```{r}

root.dir <- "data/main/"
sweepID <- "N_22_w_1p6_m_0.00005_rep_08"
dir <- paste0(root.dir,sweepID,"/")

z <- readRDS("tests/data/pearsonR.Rds")
z <- z[z$id1==sweepID,]
z <- z[grepl("_10_",z$id2),]

id <- z$id2[1]
ntp <- as.numeric(unlist(strsplit(id,"_"))[4])

x <- proc_sim(paste0(dir,"train/00000/"),seq(0,2000,length.out=ntp))

misseg_rate<- 0.00005
min_obs=10
min_tp=0

fit_obj <- readRDS(paste0(dir,"sweep_fits/",id,".Rds"))

x0 <- opt_g_free(x,min_obs,min_tp)
  x1 <- wrap_neighbor_fitness(x,x0,pm0=misseg_rate)
  x0$id <- "fq"
  x1$id <- "nn"
  xo <- rbind(x0,x1)
```

