---
title: "characterizing number of paths"
author: "Richard J Beck"
date: "`r Sys.Date()`"
output:
  md_document:
    variant: markdown_github
---
**Characterization of fitness landscapes**
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir="~/projects/008_birthrateLandscape/ALFA-k")
```

```{r}
library(fields)
source("utils/ALFA-K.R")
library(ggplot2)
```

Some ideas & thoughts

Do any CNAs ever reach fixation?

Quantifying how many distinct karyotypes appear over N simulations? Or how many distinct CNA's?

Quantifying the number of monotonically increasing pathways starting from the diploid state? This would be a good estimate of number of evolutionary paths when mis-segregation rate is low. 

Perhaps the best thing to to would be to convert the region where the landscape is valid to a network and analyse that?

```{r}
library(Matrix)
x <- readRDS("example_data/hTERTcomb_fit.Rds")
fit <- x$fit
k <- unlist(lapply(x$xo,function(xxi) rownames(xxi)[xxi$id=="fq"]))
k <- unique(k)

k1 <- gen_all_neighbours(k)
k1 <- apply(k1,1,paste,collapse=".")

k2 <- gen_all_neighbours(k1)
k2 <- apply(k2,1,paste,collapse=".")



```

```{r}

x0 <- matrix(rep(2,22),nrow=1)
x1 <- gen_all_neighbours(apply(x0,1,paste,collapse="."))
x2 <- gen_all_neighbours(apply(x1,1,paste,collapse="."))

expand_paths <- function(xn_1){
  x0 <- tail(xn_1,1)
  xn <- gen_all_neighbours(x0)
  fn_1 <-c(predict(fit,matrix(s2v(x0),nrow=1)))
  fn <- c(predict(fit,xn))
  if(sum(fn>fn_1)<1) return(NULL)
  xn <- xn[fn>fn_1,,drop=F]
  xn <- lapply(1:nrow(xn), function(i){
    c(xn_1,paste(xn[i,],collapse="."))
  })
}

wrap_paths <- function(k){
  x0 <- list(k)
  x1 <- unlist(lapply(x0,expand_paths),recursive=F)
  x2 <- unlist(lapply(x1,expand_paths),recursive=F)
  c(length(x0),length(x1),length(x2))
}

x <- readRDS("example_data/hTERTcomb_fit.Rds")
fit <- x$fit
df <- data.frame(karyotype=unlist(lapply(x$xo,function(xxi) rownames(xxi)[xxi$id=="fq"])),
                 f_est = unlist(lapply(x$xo,function(xxi) xxi$f_est[xxi$id=="fq"])))
df <- df[!duplicated(df$karyotype),]

df2 <- do.call(rbind,pbapply::pblapply(df$k,wrap_paths))
df2 <- data.frame(df2)

df <- cbind(df,df2)
df$ploidy <- sapply(df$karyotype,function(k){
  mean(s2v(k))
})

p <- ggplot(df,aes(x=f_est,y=X3))+
  geom_point()
p

p <- ggplot(df,aes(x=ploidy,y=X3))+
  geom_point()
p

```
