---
title: "ploidy_roughness_relationship"
author: "Richard J Beck"
date: "2023-06-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir="~/projects/008_birthrateLandscape/ALFA-k")
```

```{r}

source("~/projects/008_birthrateLandscape/karyotype_evolution/rscripts/analysis_functions_v2.R")
library(ggplot2)
library(fields)

```

ROUGHNESS METRIC APPLIED TO HTERT LANDSCAPE:

```{r}

roughness_meas <- function(fit,karyotypes){
  
  xmat <- do.call(rbind,lapply(karyotypes, function(i) as.numeric(unlist(strsplit(i,split="[.]")))))
  
  nn <- lapply(karyotypes,gen_all_neighbours)
  f0 <- predict(fit,xmat)
  roughness <- sapply(1:length(nn), function(i){
    f0 <- f0[i]
    fn <- predict(fit,nn[[i]])
    mean(abs(f0-fn))
  })
  
  ploidy <- apply(xmat,1,mean)
  
  data.frame(ploidy=ploidy,roughness=roughness)
}

x <- readRDS("data/cellLines/03_alfak/hTERTcombined.Rds")

karyotypes <- unique(unlist(lapply(x$xo,rownames)))
fit <- x$fit
df <- roughness_meas(fit,karyotypes)
```

MORAN METRIC ON HTERT COMBINED LANDSCAPE
```{r}

moran_freq <- function(karyotypes,fit,m=2){
  
  xmat <- do.call(rbind,lapply(karyotypes, function(i){
    as.numeric(unlist(strsplit(i,split="[.]")))
  }))
  fitness <- predict(fit,xmat)
  ploidy <- apply(xmat,1,mean)
  xbar <- mean(fitness)
  N <- length(fitness)
  d <- as.matrix(dist(xmat))
  w <- (1/d^m)#d*(d==1)
  diag(w) <- 0
  w <- apply(w,1,function(wi) wi/sum(wi))
  zi <- fitness-xbar
  
  m2 <- sum(zi^2)/N
  
  Ii <- sapply(1:length(fitness), function(i){
    zi[i]/m2*sum(w[i,]*zi)
  })  
  data.frame(Ii,ploidy)
}

x <- readRDS("data/cellLines/03_alfak/hTERTcombined.Rds")

karyotypes <- unique(unlist(lapply(x$xo,rownames)))
fit <- x$fit
df <- moran_freq(karyotypes,fit)
plot(df$ploidy,df$Ii)
```


SCRIPT THAT COMPUTES THE ROUGHNESS METRICS FOR EACH FREQUENT KARYOTYPE.
Fit the Krig models:
```{r,eval=F}

fit_krig <- function(fi,dirin,dirout){
  wl <- unlist(strsplit(fi,split="_"))[4]
  
  x2 <- readRDS(paste0(dirin,fi))
  xmat <- do.call(rbind,lapply(rownames(x2), function(i){
    as.numeric(unlist(strsplit(i,split="[.]")))
  }))
  fit <- Krig(xmat,x2$f_est)
  saveRDS(fit,paste0(dirout,fi))
}



dirout <- "proc_data/03_fitted_krigs/"
dirin <- "proc_data/02_testing_nn_inference/output/gfree/"

ff <- list.files(dir2)

x <- pbapply::pblapply(ff,fit_krig,dirin=dirin,dirout=dirout)

```

Run the landscape metrics stats:


```{r,eval=F}

source("rscripts/acf_stats.R")
source("rscripts/roughness_stats.R")

```



Function to filter out non working sims:

```{r}


dir <- "proc_data/04_testing_spline_validation_nn/gfree/"
ff <- list.files(dir)

get_R2 <- function(fi){
  fname <- unlist(strsplit(fi, split="[.]"))[1]
  x <- readRDS(paste0(dir,fi))
  rsq <- R2(x$yclone,x$ykrig)
  data.frame(fname,rsq)
}

r2df <- do.call(rbind,lapply(ff,get_R2))
r2df <- r2df[r2df$rsq>0.3,]

```

Show how the statistics look:
```{r}

m <- readRDS("proc_data/05_landscape_metrics/moran_ABM_fq.Rds")
r <- readRDS("proc_data/05_landscape_metrics/roughness_ABM_fq.Rds")
r <- r[r$sim%in%r2df$fname,]

r$wl <- gsub("p",".",r$wl)
m$wl <- gsub("p",".",m$wl)

p_moran <- ggplot(m,aes(x=wl,y=Ii))+
  geom_violin()+
  scale_x_discrete(expression(lambda))+
  scale_y_continuous("local moran's i")+
  theme_classic(base_size=8)
p_moran

p_rough <- ggplot(r,aes(x=wl,y=roughness))+
  geom_violin()+
  scale_x_discrete(expression(lambda))+
  scale_y_continuous("ruggedness",limits=c(0,0.5))+
  theme_classic(base_size=8)
p_rough
```

```{r}

comparison_heatmap <- function(x,za2=1.96,zb=0.84){
  Za2 <- 1.96
  Zb <- 0.84
  x <- split(x,f=x$wl)

  n <- do.call(rbind,lapply(x, function(xi){
    sapply(x, function(xj){
      i <- xi$Ii
      j <- xj$Ii
      sigma <- sd(c(i-mean(i),j-mean(j)))
      delta <- mean(i)-mean(j)
      2*(Za2+Zb)*(sigma/delta)^2
    })
  }))
  return(n)
}

merged_comparison_heatmap <- function(x1,x2,za2=1.96,zb=0.84,lab=NaN){
  r21 <- nrow(x2)/nrow(x1)
  n1 <- comparison_heatmap(x1,za2,zb)
  n1 <- cbind(rep(Inf,4),n1)
  n1 <- rbind(rep(Inf,5),n1)
  rownames(n1)[1] <- "0p1"
  colnames(n1)[1] <- "0p1"
  n1 <- n1*r21
  n2 <- comparison_heatmap(x2,za2,zb)
  n2[upper.tri(n2)]<-n1[upper.tri(n1)]
  
  n <- reshape2::melt(n2)
  n$Var1 <- as.character(n$Var1)
  n$Var2 <- as.character(n$Var2)
  n
}

dir <- "proc_data/05_landscape_metrics/"

x1 <- readRDS("proc_data/05_landscape_metrics/moran_ABM_nn.Rds")
x1 <- x1[x1$sim%in%r2df$fname,]
x2 <- readRDS("proc_data/05_landscape_metrics/moran_ABM_fq.Rds")
 
#x0 <- comparison_heatmap(x0,lab="direct")
#x1 <- comparison_heatmap(x1,lab="ABM nn")



#x2 <- comparison_heatmap(x2,lab="ABM fq")

#x <- rbind(x1,x2)
x <- merged_comparison_heatmap(x1,x2)

x$Var1 <- gsub("p",".",x$Var1)
x$Var2 <- gsub("p",".",x$Var2)
x$toplab <- "frequent + neighbours"
x$sidelab <- "frequent only"

p_moran_ABM <- ggplot(x,aes(x=Var1,y=Var2,fill=value))+
  facet_grid(cols=vars(toplab),rows=vars(sidelab))+
  geom_raster()+
  scale_fill_viridis_c("observations\nneeded for \nsignificance",trans="log",
                        label = function(x) sprintf("%.1f", x))+
  scale_x_discrete(expression(lambda~(landscape~B)))+
  scale_y_discrete(expression(lambda~(landscape~A)))+
  theme_classic(base_size=8)+
    geom_abline(color="red")+
  theme(legend.key.height = unit(5,"mm"),
        legend.key.width = unit(3,"mm"))
p_moran_ABM



```


```{r}

comparison_heatmap <- function(x,za2=1.96,zb=0.84,lab=NaN){
  Za2 <- 1.96
  Zb <- 0.84
  x <- split(x,f=x$wl)

  n <- do.call(rbind,lapply(x, function(xi){
    sapply(x, function(xj){
      i <- xi$roughness
      j <- xj$roughness
      sigma <- sd(c(i-mean(i),j-mean(j)))
      delta <- mean(i)-mean(j)
      2*(Za2+Zb)*(sigma/delta)^2
    })
  }))
  return(n)
}
merged_comparison_heatmap <- function(x1,x2,za2=1.96,zb=0.84,lab=NaN){  
  r21 <- nrow(x2)/nrow(x1)
  n1 <- comparison_heatmap(x1,za2,zb)
  n1 <- n1*r21
  n2 <- comparison_heatmap(x2,za2,zb)
  n2[upper.tri(n2)]<-n1[upper.tri(n1)]
  n <- reshape2::melt(n2)
  n$Var1 <- as.character(n$Var1)
  n$Var2 <- as.character(n$Var2)
  n
}

dir <- "proc_data/05_landscape_metrics/"

x1 <- readRDS("proc_data/05_landscape_metrics/roughness_ABM_nn.Rds")
x1 <- x1[x1$sim%in%r2df$fname,]
x2 <- readRDS("proc_data/05_landscape_metrics/roughness_ABM_fq.Rds")
x2 <- x2[x2$sim%in%r2df$fname,]
x <- merged_comparison_heatmap(x1,x2)

x$Var1 <- gsub("p",".",x$Var1)
x$Var2 <- gsub("p",".",x$Var2)
x$toplab <- "frequent + neighbours"
x$sidelab <- "frequent only"

p_rough_ABM <- ggplot(x,aes(x=Var1,y=Var2,fill=value))+
  facet_grid(cols=vars(toplab),rows=vars(sidelab))+
  geom_raster()+
  scale_fill_viridis_c("observations\nneeded for \nsignificance",trans="log",
                        label = function(x) sprintf("%.1f", x))+
  scale_x_discrete(expression(lambda~(landscape~B)))+
  scale_y_discrete(expression(lambda~(landscape~A)))+
  theme_classic(base_size=8)+
  geom_abline(color="red")+
  theme(legend.key.height = unit(5,"mm"),
        legend.key.width = unit(3,"mm"))
p_rough_ABM


```

Show the results for the cell lines

```{r}

x1 <- readRDS("proc_data/05_landscape_metrics/moran_hTERT_fq.Rds")
x2 <- readRDS("proc_data/05_landscape_metrics/roughness_hTERT_fq.Rds")

x1$rploidy <- round(x1$ploidy)
x2$rploidy <- round(x2$ploidy)
t.test(x1$Ii[x1$rploidy==2],x1$Ii[x1$rploidy==4])
t.test(x2$roughness[x2$rploidy==2],x2$roughness[x2$rploidy==4])

p_moran_htert <- ggplot(x1,aes(x=ploidy,y=Ii,color=cellLine))+
  geom_point()+
  scale_x_continuous("ploidy")+
  scale_y_continuous("local Moran's i")+
  scale_color_discrete("")+
  theme_classic(base_size=8)+
  theme(legend.position = "top",legend.key.size = unit(5,"mm"))
p_moran_htert

p_rough_htert <- ggplot(x2,aes(x=ploidy,y=roughness,color=cellLine))+
  geom_point()+
  scale_x_continuous("ploidy")+
  scale_y_continuous("ruggedness")+
  scale_color_discrete("")+
  theme_classic(base_size=8)+
  theme(legend.position = "top",legend.key.size = unit(5,"mm"))
p_rough_htert


```


```{r}

p <- cowplot::plot_grid(p_moran,p_rough,
                        p_moran_ABM,p_rough_ABM,
                   p_moran_htert,p_rough_htert,
                   rel_widths = c(2.7,4,3),
                   nrow = 2,labels=c("A","C","E","B","D","F"),byrow = F)


ggsave("figures/ploidy_roughness_relationship.png",plot=p,width=7.3,height=4,units = "in")

```

```{r}

comparison_heatmap <- function(x,za2=1.96,zb=0.84,lab=NaN){
  Za2 <- 1.96
  Zb <- 0.84
  
  x <- split(x,f=x$wl)

  n <- do.call(rbind,lapply(x, function(xi){
    sapply(x, function(xj){
      i <- xi$roughness
      j <- xj$roughness
      sigma <- sd(c(i-mean(i),j-mean(j)))
      delta <- mean(i)-mean(j)
      2*(Za2+Zb)*(sigma/delta)^2
    })
  }))
  
  n <- reshape2::melt(n)
  n$Var1 <- as.character(n$Var1)
  n$Var2 <- as.character(n$Var2)
  n$lab <- lab

  n
}

x1 <- readRDS("proc_data/05_landscape_metrics/roughness_ABM_nn.Rds")
x1 <- x1[x1$sim%in%r2df$fname,]
x2 <- readRDS("proc_data/05_landscape_metrics/roughness_ABM_fq.Rds")
x2 <- x2[x2$sim%in%r2df$fname,]
x1 <- comparison_heatmap(x1,lab="ABM nn")
x2 <- comparison_heatmap(x2,lab="ABM fq")

x <- rbind(x1,x2)
x$Var1 <- gsub("p",".",x$Var1)
x$Var2 <- gsub("p",".",x$Var2)



p_rough_ABM_nn <- ggplot(x[x$lab=="ABM nn",],aes(x=Var1,y=Var2,fill=value/44))+
  geom_raster()+
  scale_fill_viridis_c("observations\nneeded for \nsignificance",trans="log")+
  scale_x_discrete("GRF wavelength (landscape B)")+
  scale_y_discrete("GRF wavelength (landscape A)")
p_rough_ABM_nn

p_rough_ABM_fq <- ggplot(x[x$lab=="ABM fq",],aes(x=Var1,y=Var2,fill=value))+
  geom_raster()+
  scale_fill_viridis_c("observations\nneeded for \nsignificance",trans="log")+
  scale_x_discrete("GRF wavelength (landscape B)")+
  scale_y_discrete("GRF wavelength (landscape A)")
p_rough_ABM_fq




```

check correlations:

```{r}

x1 <- readRDS("proc_data/05_landscape_metrics/moran_ABM_fq.Rds")
x2 <- readRDS("proc_data/05_landscape_metrics/roughness_abm_fq.Rds")

x <- data.frame(wl=x1$wl,Ii=x1$Ii,r=x2$roughness)

p <- ggplot(x,aes(x=Ii,y=r,color=wl))+
  geom_point(alpha=0.1)
p

```



Restricted to cases where the landscape estimation worked we do see differences in roughness metric. Roughness metric appears to follow a lognormal distribution. Based on krigs that worked we estimated how many samples we'd need to register statistically significant difference in roughness between various fitness landscapes. Answer = up to 100 depending on how different the landscapes are and desired level of significance. 

```{r}

assess_ruggedness <- function(fi,dir0,dir2,dir3){
  wl <- unlist(strsplit(fi,split="_"))[4]
  
  xo <- readRDS(paste0(dir0,fi))
  x0mat <- do.call(rbind,lapply(rownames(xo), function(i){
    as.numeric(unlist(strsplit(i,split="[.]")))
  }))
  x2 <- readRDS(paste0(dir2,fi))
  xmat <- do.call(rbind,lapply(rownames(x2), function(i){
    as.numeric(unlist(strsplit(i,split="[.]")))
  }))
   fit <- readRDS(paste0(dir3,fi))
  
  roughness <- sapply(1:nrow(xo), function(i){
    n <- gen_all_neighbours(rownames(xo)[i])
    fn <- c(predict(fit,n))
    f0 <- c(predict(fit,matrix(x0mat[i,],nrow=1)))
    mean(abs(f0-fn))  
  })
  
  data.frame(roughness,wl)
}

checkR2 <- function(xpath){
  x <- readRDS(xpath)
  ykrig <- x$ykrig - mean(x$yclone)
  yclone <- x$yclone-mean(x$yclone)
  ytru <- x$ytru-mean(x$ytru)
  
  max(-1,R2(obs=yclone,pred=ykrig))
}


df <- readRDS("~/projects/008_birthrateLandscape/karyotype_evolution/proc_data/03_testing_spline_inference_nnin.Rds")
dff <- paste0("N_",df$nchrom,"_w_",df$wl,"_rep_",df$rep,".Rds") 
dff <- dff[df$R2>0.2]
dir0 <- "proc_data/01_testing_frequent_clone_inference/output/gfree/"
dir2 <- "proc_data/02_testing_nn_inference/output/gfree/"
dir3 <- "proc_data/03_fitted_krigs/"
dir4 <- "proc_data/04_testing_spline_validation/gfree/"
f4 <- list.files(dir4)
R4 <- sapply(paste0(dir4,f4),checkR2)
f4 <- f4[R4>0]
ff <- list.files(dir2)
ff <- ff[ff%in%dff]
x <- pbapply::pblapply(ff,assess_ruggedness,dir0=dir0,dir2=dir2,dir3=dir3)
y <- do.call(rbind,x)
z <- do.call(rbind,lapply(x, function(xi) data.frame(wl=xi$wl[1],roughness=xi$roughness,n=nrow(xi))))

p <- ggplot(y,aes(x=wl,y=roughness))+
  geom_jitter()
p

p <- ggplot(z,aes(x=wl,y=roughness,color=n))+
  geom_jitter()
p

p <- ggplot(z,aes(x=wl,y=roughness))+
  geom_violin()+
  geom_jitter()+
  scale_y_log10()
p

x <- split(y,f=y$wl)


Za2 <- 1.96
Zb <- 0.84

n <- do.call(rbind,lapply(x, function(xi){
  sapply(x, function(xj){
    i <- log(xi$roughness)
    j <- log(xj$roughness)
    sigma <- sd(c(i-mean(i),j-mean(j)))
    delta <- mean(i)-mean(j)
    2*(Za2+Zb)*(sigma/delta)^2
  })
}))

ceiling(n)

n <- reshape2::melt(n)

p <- ggplot(n,aes(x=Var1,y=Var2,fill=value))+
  geom_raster()+
  scale_fill_viridis_c("observations\nneeded for \nsignificance")+
  scale_x_discrete("GRF wavelength (landscape B)")+
  scale_y_discrete("GRF wavelength (landscape A)")+
  ggtitle("roughness metric")
p

```
We do not have enough information to establish significance of different ploidies:

```{r}

x <- readRDS("proc_data/moran_freq_kary.Rds")
x <- split(x,f=x$wl)


Za2 <- 1.96
Zb <- 0.84

n <- do.call(rbind,lapply(x, function(xi){
  sapply(x, function(xj){
    i <- xi$Ii
    j <- xj$Ii
    sigma <- sd(c(i-mean(i),j-mean(j)))
    delta <- mean(i)-mean(j)
    2*(Za2+Zb)*(sigma/delta)^2
  })
}))

ceiling(n)

n <- reshape2::melt(n)

p <- ggplot(n,aes(x=Var1,y=Var2,fill=value))+
  geom_raster()+
  scale_fill_viridis_c("observations\nneeded for \nsignificance",trans="log",breaks=c(1,3,9,27,81))+
  scale_x_discrete("GRF wavelength (landscape B)")+
  scale_y_discrete("GRF wavelength (landscape A)")+
  ggtitle("autocorrelation metric")
p

```
```{r}

roughness_meas <- function(cellLine){
  m <- readRDS(paste0("salehi_data/03_inference_output/",cellLine,"/krig.Rds"))
  x <- readRDS(paste0("salehi_data/03_inference_output/",cellLine,"/frequent_clones.Rds"))
  xmat <- do.call(rbind,lapply(rownames(x), function(i) as.numeric(unlist(strsplit(i,split="[.]")))))
  
  nn <- lapply(rownames(x),gen_all_neighbours)
  f0 <- predict(m,xmat)
  roughness <- sapply(1:length(nn), function(i){
    f0 <- f0[i]
    fn <- predict(m,nn[[i]])
    mean(abs(f0-fn))
  })
  
  ploidy <- apply(xmat,1,mean)
  
  data.frame(ploidy=ploidy,roughness=roughness,cellLine=cellLine)
}

df <- rbind(roughness_meas("hTERTa"),roughness_meas("hTERTb"))#,
#            roughness_meas("SA1035"),roughness_meas("SA535"),roughness_meas("SA609ln1"))

p <- ggplot(df,aes(x=ploidy,y=roughness,color=cellLine))+
  geom_point()+
  scale_y_log10()
p

y1 <- df$roughness[round(df$ploidy)==2]
y2 <- df$roughness[round(df$ploidy)==4]

t.test(log(y1),log(y2))
```

Can we assess roughness without the fitness landscape?

For rough landscapes features should include large jumps in fitness, fewer unique clones. 
For smooth landscapes there should be smaller jumps in fitness and less unique clones. 

```{r}
dir <- "proc_data/01_testing_frequent_clone_inference/output/gfree/"
ff <- list.files(dir)

x <- do.call(rbind,lapply(ff, function(fi){
  x <- readRDS(paste0(dir,fi))
wl <- unlist(strsplit(fi,split="_"))[4]
wl <- as.numeric(gsub("p",".",wl))
xstats <- data.frame(wl=wl,Nf=nrow(x),
                     rangef = max(x$f_est)-min(x$f_est),
                     sdn = sd(x$n),
                     meann = mean(x$n),
                     mediann = median(x$n))
return(xstats)
}))
x$sdn[is.na(x$sdn)] <- 0
corrplot::corrplot(cor(x))

p <- ggplot(x,aes(x=wl,y=Nf))+
  geom_point()
p

```
We are trying to see whether close neighbors have large or small fitness jumps
```{r}

dir <- "proc_data/01_testing_frequent_clone_inference/output/gfree/"
ff <- list.files(dir)

get_stats <- function(fi){
  wl <- unlist(strsplit(fi,split="_"))[4]
wl <- as.numeric(gsub("p",".",wl))
  x <- readRDS(paste0(dir,fi))
x <- x[order(x$f_est),]
xmat <- do.call(rbind,lapply(rownames(x), function(xi){
  as.numeric(unlist(strsplit(xi,split="[.]")))
}))

d <- as.matrix(dist(xmat))

mj <- do.call(rbind,lapply(1:(nrow(d)-1), function(i){
  di <- d[i,(i+1):ncol(d)]
  nn <- sum(di==min(di))
  mdf <- mean(abs(x$f_est[i]-x$f_est[which(di==min(di))+i]))
  data.frame(nn,mdf)
}))

xstats <- data.frame(wl=wl,mdf = mean(mj$mdf), mnn = mean(mj$nn))
return(xstats)
}

x <- do.call(rbind,lapply(ff,get_stats))

p <- ggplot(x,aes(x=wl,y=mdf))+
  geom_point()
p

p <- ggplot(x,aes(x=wl,y=mnn))+
  geom_jitter(height=0,width=0.02)
p

p <- ggplot(x,aes(x=wl,y=mnn/mdf))+
  geom_jitter(height=0,width=0.02)
p

```
```{r}

dir <- "ABM/output/nchrom_22/"
f0 <- list.files(dir)

check_nn <- function(fi){
  wl <- unlist(strsplit(fi,split="_"))[4]
  foi <- list.files(paste0(dir,fi,"/train"))[1]
  fpath <- paste0(dir,fi,"/train/",foi)
  xi <- proc_sim(fpath,times=seq(400,2800,400))$x
  n <- rowSums(xi)
  xmat <- do.call(rbind,lapply(rownames(xi), function(xx){
    as.numeric(unlist(strsplit(xx,split="[.]")))
  }))
  d <- as.matrix(dist(xmat))
  n1 <- apply(d,1, function(di) sum(di==1))
  nr2 <- apply(d,1, function(di) sum((di>1)&(di<2)))
  n2 <- apply(d,1, function(di) sum(di==2))
  data.frame(wl,mn1=mean(n1/n),mn2=mean(n2/n),mnr2 = mean(nr2/n))
}

df <- do.call(rbind,pbapply::pblapply(f0, check_nn))

p <- ggplot(df,aes(x=wl,y=mn1))+
  geom_point()
p

p <- ggplot(df,aes(x=wl,y=mn2))+
  geom_point()
p


p <- ggplot(df,aes(x=wl,y=mnr2))+
  geom_point()
p


```

```{r}

x <- readRDS("proc_data/roughness_stats.Rds")

p <- ggplot(x,aes(x=n,y=n2,color=wl))+
  facet_wrap(~wl)+
  geom_density2d()+
  scale_x_log10()+
  scale_y_log10()
p

y <- x[,-2]
y$wl <- as.numeric(gsub("p",".",x$wl))
z <- reshape2::melt(x,id.vars="wl")
p <- ggplot(z,aes(x=wl,y=value))+
  facet_wrap(~variable,scales="free")+
  geom_violin()+
  scale_y_log10()
p

p <- ggplot(z,aes(x=wl,y=value))+
  facet_wrap(~variable,scales="free")+
  geom_violin()+
  scale_y_continuous()
p
corrplot::corrplot(cor(y))
```
```{r}

library(caret)
x <- readRDS("proc_data/roughness_stats.Rds")
x$wl <- as.numeric(gsub("p",".",x$wl))
x <- x[,c("wl","n","n2")]
for(i in 1:ncol(x)) {
  x[,i] <- x[,i]+0.1
  x[,i] <- log(x[,i])
}
fitControl <- trainControl(method = "repeatedcv",   
                           number = 10,     # number of folds
                           repeats = 10)

model <- train(wl ~ .,
               data = x,
               method = "lm",
               trControl = fitControl)
pred <- predict(model,x)

dfp <- data.frame(wl=as.character(exp(x$wl)),pred_wl=exp(pred))

plot(exp(x$wl)-0.1,exp(pred)-0.1)

p <- ggplot(dfp,aes(x=wl,y=pred_wl))+
  geom_violin()+
  geom_jitter()
p

```

```{r}

x <- readRDS("proc_data/indi_xstats.Rds")
nrow(x)
z <- reshape2::melt(x,id.vars="wl")
p <- ggplot(z,aes(x=wl,y=value))+
  facet_wrap(~variable,scales="free")+
  geom_jitter(height=0,width=0.2,alpha=0.1)+
  scale_y_log10()
p

x <- readRDS("proc_data/indi_xstats2.Rds")
nrow(x)
#x <- x[x$n>5,]
z <- reshape2::melt(x,id.vars="wl")
p <- ggplot(z,aes(x=wl,y=value))+
  facet_wrap(~variable,scales="free")+
  geom_violin()+
  scale_y_log10()
p

p <- ggplot(z,aes(x=wl,y=value))+
  facet_wrap(~variable,scales="free")+
  geom_violin()
p



```

```{r}
library(factoextra)
library(cluster)
x <- readRDS("proc_data/xstats.Rds")
z <- reshape2::melt(x,id.vars="wl")
p <- ggplot(z,aes(x=wl,y=value))+
  facet_wrap(~variable,scales="free")+
  geom_point()
p
id <- x[,1]
x <- x[,-1]
x <- scale(x)

fviz_nbclust(x, kmeans, method = "wss")

km <- kmeans(x, centers = 5, nstart = 25)

tsne <- Rtsne::Rtsne(x)
y <- data.frame(tsne$Y)
y$id <- id
y$cluster <- km$cluster

p <- ggplot(y,aes(x=X1,y=X2,color=id))+
  geom_point()
p

p <- ggplot(y,aes(x=X1,y=X2,color=cluster))+
  geom_point()
p

p <- ggplot(y,aes(x=id,y=cluster))+
  geom_jitter()
p

```


```{r}
library(caret)
x <- readRDS("proc_data/xstats.Rds")
x$wl <- as.numeric(gsub("p",".",x$wl))
fitControl <- trainControl(method = "repeatedcv",   
                           number = 10,     # number of folds
                           repeats = 10)

model <- train(wl ~ .,
               data = x,
               method = "lasso",
               trControl = fitControl)
pred <- predict(model,x)

plot(x$wl,pred)

```

```{r}

 x <- readRDS("salehi_data/03_inference_output/hTERTa/frequent_clones.Rds")

```
```{r}
assess_acf <- function(fi,dir0,dir2,dir3){
  wl <- unlist(strsplit(fi,split="_"))[4]
  
  xo <- readRDS(paste0(dir0,fi))
  x0mat <- do.call(rbind,lapply(rownames(xo), function(i){
    as.numeric(unlist(strsplit(i,split="[.]")))
  }))
  x2 <- readRDS(paste0(dir2,fi))
  xmat <- do.call(rbind,lapply(rownames(x2), function(i){
    as.numeric(unlist(strsplit(i,split="[.]")))
  }))
  fit <- readRDS(paste0(dir3,fi))
  d <- as.matrix(dist(xmat))
  d <- d[(1:nrow(xo)),]
  
  acf <- do.call(rbind,lapply(1:nrow(d), function(i){
    data.frame(d=d[i,],f0=xo$f_est[i],f2=x2$f_est)
  }))
  
  acf$wl <- wl
  acf
}


dir0 <- "proc_data/01_testing_frequent_clone_inference/output/gfree/"
dir2 <- "proc_data/02_testing_nn_inference/output/gfree/"
dir3 <- "proc_data/03_fitted_krigs/"


dir4 <- "proc_data/04_testing_spline_validation/gfree/"
f4 <- list.files(dir4)
R4 <- sapply(paste0(dir4,f4),checkR2)
f4 <- f4[R4>0]
ff <- list.files(dir2)
ff <- ff[ff%in%f4]

x <- pbapply::pblapply(ff,assess_acf,dir0=dir0,dir2=dir2,dir3=dir3)
y <- do.call(rbind,x)

y <- split(y,f=interaction(y$d,y$wl))

z <- do.call(rbind,lapply(y,function(yi){
  mn <- mean(yi$f0)
  vacf <- sum((mn-yi$f2)*(mn-yi$f0))/sum((mn-yi$f0)^2)
  data.frame(d=yi$d[1],wl=yi$wl[1],acf=vacf,n=nrow(yi))
}))

z <- z[z$d<3,]

p <- ggplot(z,aes(x=d,y=acf,color=wl))+
  geom_smooth()
p

p <- ggplot(z,aes(x=d,y=acf,color=wl))+
  geom_point()
p
```

```{r}

x1 <- readRDS("salehi_data/02_optim_data/hTERTa.Rds")$x
x2 <- readRDS("salehi_data/02_optim_data/hTERTb.Rds")$x

x <- x1

xmat <- do.call(rbind,lapply(rownames(x), function(i) as.numeric(unlist(strsplit(i,split="[.]")))))

d <- as.matrix(dist(xmat))
d <- d==1

ploidy <- apply(xmat,1,mean)

df <- do.call(rbind,lapply(1:ncol(x), function(i){
  df <- data.frame(ploidy,nn=d[,i]*(x[,i]>0), nntot=d[,i]*x[,i], n0 = x[,i],t=i)
}))



p <- ggplot(df,aes(x=n0,y=nntot,color=ploidy))+
  geom_point()+
  scale_color_viridis_c()
p

pca <- data.frame(prcomp(xmat)$x[,1:2])
pca$ploidy <- ploidy

p <- ggplot(pca,aes(x=PC1,y=PC2,color=ploidy))+
  geom_point()
p

```
```{r}

m <- readRDS("salehi_data/03_inference_output/hTERTa/krig.Rds")
x <- readRDS("salehi_data/03_inference_output/hTERTa/frequent_clones.Rds")
xmat <- do.call(rbind,lapply(rownames(x), function(i) as.numeric(unlist(strsplit(i,split="[.]")))))

nn <- lapply(rownames(x),gen_all_neighbours)
fsd <- sapply(nn, function(ni) sd(predict(m,ni)))
ploidy <- apply(xmat,1,mean)

plot(ploidy,fsd)
```


```{r}

x <- readRDS("salehi_data/03_inference_output/hTERTa/frequent_clones.Rds")
y <- readRDS("salehi_data/02_optim_data/hTERTa.Rds")$x
y <- y[rownames(x),]

m <- do.call(rbind,lapply(rownames(x), function(xi) as.numeric(unlist(strsplit(xi,split="[.]")))))

ploidy <- apply(m,1,median)

m2 <- m[ploidy==2,]
m4 <- m[ploidy==4,]

```

```{r}
ma <- readRDS("salehi_data/03_inference_output/hTERTa/krig.Rds")
xa <- readRDS("salehi_data/03_inference_output/hTERTa/nn_clones.Rds")

mb <- readRDS("salehi_data/03_inference_output/hTERTb/krig.Rds")
xb <- readRDS("salehi_data/03_inference_output/hTERTb/nn_clones.Rds")

x <- rownames(xa)[rownames(xa)%in%rownames(xb)]

x <- data.frame(fa=xa[x,"f_est"],fb=xb[x,"f_est"])
p <- ggplot(x,aes(x=fa,y=fb))+
  geom_point()+
  scale_x_continuous("fitness in hTERTa")+
  scale_y_continuous("fitness in hTERTb")
p

```