---
title: "landscape_visualisation"
author: "Richard J Beck"
date: "3/10/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir="~/projects/008_birthrateLandscape/ALFA-K/")
```

Input data for this sweep was generated using the following command:

Rscript generate_ABM_sweep.R -n data/misrate -r 100 -w 0.8 -m 0.00001,0.00003,0.0001,0.0003,0.001 
Rscript process_abm_sweep.R -n data/misrate 
Rscript sweep_summary_stats -n data/misrate -o data/misrate_stats.Rds

```{r}
source("~/projects/008_birthrateLandscape/karyotype_evolution/rscripts/landscape_functions.R")
source("~/projects/008_birthrateLandscape/karyotype_evolution/rscripts/analysis_functions_v2.R")
library(ggplot2)
library(fields)
library(parallel)
```
Check performance vs misseg rate and look for summary statistics that explain good performance

```{r}
x <- readRDS("data/misrate_stats.Rds")
df <- reshape2::melt(x[,c("m","r_fq","r_nn","r_d2n")],id.vars="m")
p <- ggplot(df,aes(x=factor(m),y=value))+
  facet_grid(rows=vars(variable))+
  geom_violin()
p

p <- ggplot(x,aes(x=nfq,y=r_fq))+
  facet_grid(rows=vars(m),scales="free")+
  geom_point()+
  geom_smooth(method="lm")
p

xc <- x[,-c(1,2,4)]
for(i in 1:ncol(xc)) xc[,i] <- as.numeric(xc[,i])
corrplot::corrplot(cor(xc,method = "spearman"))
xc <- split(xc,f=xc$m)

for(xci in xc){
  corrplot::corrplot(cor(xci,method = "spearman"))
}

```

Then all sims were processed and landscapes estimated (this does take quite a while):
```{r,eval=F}
source("rscripts/process_abm_sims.R")
```

Following shows accuracy of landscape predictions as a function of mis-segregation rate:

```{r}
get_sweep_results <- function(fi){
  if(length(list.files(paste0(dir,fi)))==0) return(NULL)
  x <- readRDS(paste0(dir,fi,"/xo.Rds"))
  xi <- split(x,f=x$id)
  res <- data.frame(t(sapply(xi, function(xij) cor(xij$f_est,xij$f_tru))))
  res$misrate <- as.numeric(unlist(strsplit(fi,split="_"))[6])
  res$nfq <- nrow(xi$fq)
  res$ntot <- nrow(x)
  return(res)
}
dir <- "proc_data/misrate/"
ff <- list.files(dir)
x <- do.call(rbind,lapply(ff, get_sweep_results))
y <- x[,c("misrate","nfq","ntot")]
x <- x[,c("misrate","fq","nn","d2n")]
x <- reshape2::melt(x,id.vars=c("misrate"))
p <- ggplot(x,aes(x=factor(misrate),y=value))+
  facet_grid(rows=vars(variable))+
  geom_violin()
p

df <- aggregate(list(value=x$value),by=list(misrate=x$misrate,variable=x$variable),median,na.rm=T)

p <- ggplot(df,aes(x=factor(misrate),y=value,color=variable))+
  geom_point()+
  scale_y_continuous("estimated fitness correlation")+
  scale_x_discrete("misseg. rate")+
  scale_color_discrete("distance to\nobserved\nkaryotype",labels=c(0,1,2))
p


p <- ggplot(y,aes(x=factor(misrate),y=nfq))+
  geom_violin()+
  scale_y_continuous("num. clones")+
  scale_x_discrete("misseg.rate")
p

p <- ggplot(y,aes(x=factor(misrate),y=ntot))+
  geom_violin()+
  scale_y_continuous("karyotypes estimated")+
  scale_x_discrete("misseg.rate")
p

```

Another question is how misrate influences exploration of fitness landscapes. Some basic metrics regarding this are calculated and presented below.

```{r}

dproc <- function(fi,dir){
  p <- as.numeric(unlist(strsplit(fi,split="_"))[6])
  r <- as.numeric(unlist(strsplit(fi,split="_"))[8])
  di <- paste0(dir,fi,"/train/00000/")
  x <- proc_sim(di,times=seq(0,3000,500))
  popf <- x$pop.fitness
  x <- x$x
  fdip <- x[paste(rep(2,22),collapse="."),]/colSums(x)
  nkary <- apply(x,2,function(xi) sum(xi>0))
  largest_clone <- apply(x,2,max)
  cnmat <- do.call(rbind,lapply(rownames(x), function(xi){
    as.numeric(unlist(strsplit(xi,split="[.]")))
  }))
  ncopies <- t(cnmat)%*%x
  for(i in 1:ncol(ncopies)){
    ncopies[,i]<- ncopies[,i]/sum(x[,i])
  }
  dm <- apply(ncopies,2,function(k) sqrt(sum((k-2)^2)))
  ploidy <- colMeans(ncopies)
  dx <- sapply(rownames(x), function(xi){
    k <- as.numeric(unlist(strsplit(xi,split="[.]")))
    sqrt(sum((k-2)^2))
  })
  d <- apply(x,2,function(xi){
    sum(xi*dx)/sum(xi)
  })
  data.frame(d,dm,ploidy,fdip,popf,time=names(d),p,rep=r,nkary,largest_clone)
}

dir <- "ABM/output/misrate/"
ff <- list.files(dir)
df <- do.call(rbind,pbapply::pblapply(ff,dproc,dir=dir))



p <- ggplot(df,aes(x=as.numeric(time),y=d,color=factor(p),group=interaction(p,rep)))+
  geom_line()+
  scale_color_discrete("misseg.\nrate")+
  scale_x_continuous("time")+
  scale_y_continuous("mean distance to founder")
p

p <- ggplot(df,aes(x=as.numeric(time),y=dm,color=factor(p),group=interaction(p,rep)))+
  geom_line()+
  scale_color_discrete("misseg.\nrate")+
  scale_x_continuous("time")+
  scale_y_continuous("distance centroid-founder")
p

p <- ggplot(df,aes(x=as.numeric(time),y=ploidy,color=factor(p),group=interaction(p,rep)))+
  facet_wrap(~p)+
  scale_color_discrete("misseg.\nrate")+
  geom_line()+
  scale_x_continuous("time")
p

p <- ggplot(df,aes(x=as.numeric(time),y=nkary,color=factor(p),group=interaction(p,rep)))+
  facet_wrap(~p)+
  scale_color_discrete("misseg.\nrate")+
  geom_line()+
  scale_y_log10("unique karyotypes")+
  scale_x_continuous("time")
p

p <- ggplot(df,aes(x=as.numeric(time),y=largest_clone,color=factor(p),group=interaction(p,rep)))+
  facet_wrap(~p)+
  scale_color_discrete("misseg.\nrate")+
  geom_line()+
  scale_y_log10("dominant clone size")+
  scale_x_continuous("time")
p

p <- ggplot(df,aes(x=as.numeric(time),y=fdip,color=factor(p),group=interaction(p,rep)))+
  facet_wrap(~p)+
  scale_color_discrete("misseg.\nrate")+
  geom_line()+
  scale_y_log10("diploid frequency")+
  scale_x_continuous("time")
p

p <- ggplot(df,aes(x=as.numeric(time),y=popf,color=factor(p),group=interaction(p,rep)))+
  facet_wrap(~p)+
  scale_color_discrete("misseg.\nrate")+
  geom_line()+
  scale_y_continuous("population fitness")+
  scale_x_continuous("time")
p

```

More informative metrics might be had via looking at results in a more fine grained manner. The idea would be to evaluate individual populations for things like measures of spatial clustering etc and check how that relates back to prediction error. 

```{r}
##see dev/testing_misrate_sweep.R (work in progress)
```

```{r}
setwd("~/projects/008_birthrateLandscape/karyotype_evolution/")
library(snifter)

preproc <- function(fi,dir){
  p <- as.numeric(unlist(strsplit(fi,split="_"))[6])
  r <- as.numeric(unlist(strsplit(fi,split="_"))[8])
  di <- paste0(dir,fi,"/train/00000/")
  x <- proc_sim(di,times=seq(0,2800,400))$x
  x <- x[rowSums(x)>5,,drop=F]
  
  rx <- do.call(rbind,lapply(rownames(x), function(rxi){
    as.numeric(unlist(strsplit(rxi,split="[.]")))
  }))
  x <- reshape2::melt(x)
  colnames(x) <- c("karyotype","time","N")
  x$p <- p
  x$r <- r
  x <- x[x$N>0,]
  list(cn=rx,df=x)
}

dir <- "ABM/output/misrate/"
ff <- list.files(dir)
r <- sapply(ff,function(fi) as.numeric(unlist(strsplit(fi,split="_"))[8]))
x <- pbapply::pblapply(ff[c(1+(0:5)*25,2+(0:5)*25,3+(0:5)*25)],preproc,dir=dir)

df <- do.call(rbind,lapply(x, function(xi) xi$df))
cn <- do.call(rbind,lapply(x, function(xi) xi$cn))
cn <- unique(cn)

tsne <- fitsne(cn)
rownames(tsne) <- apply(cn,1,paste,collapse=".")
colnames(tsne) <- c("tsne1","tsne2")

df <- cbind(df,tsne[df$karyotype,])


#df <- cbind(df,tsne)
df$tp <- round(df$time/500)
p <- ggplot(df,aes(x=tsne1,y=tsne2,color=factor(r),size=N))+
  facet_grid(cols=vars(p),rows=vars(tp))+
  geom_point(alpha=0.5)+
  scale_color_viridis_d()
p

```


TSNE PLOT?
```{r}

dir <- "ABM/output/misrate/"
ff <- list.files(dir)
pm <- sapply(ff, function(fi) as.numeric(unlist(strsplit(fi,split="_"))[6]))
pm <- split(pm,f=pm)

x <- do.call(rbind,lapply(pm,function(pmi){
  do.call(rbind,lapply(1:3, function(j){
    pm <- as.numeric(pmi)[j]
    fi <- names(pmi)[j]
    r <- paste("rep:",as.numeric(unlist(strsplit(fi,split="_"))[8]))
    di <- paste0(dir,fi,"/train/00000/")
    x <- proc_sim(di,times=seq(0,3000,500))$x
    x <- x[sample(1:nrow(x)),]
    
    #x<- x[order(rowSums(x),decreasing=T),]
    xtop <- x[rowSums(x)>5,]
    xbot <- x[!rowSums(x)>5,]
    xbot <- colSums(xbot)
    x <- rbind(xtop,others=xbot)
    rownames(x)[1:(nrow(x)-1)]<- 1:(nrow(x)-1)
    for(i in 1:ncol(x)) x[,i] <- x[,i]/sum(x[,i])
    
    x <- reshape2::melt(x)
    colnames(x) <- c("karyotype","time","frequency")
    x$pm<-pm
    x$rep <- r
    return(x)
  }))
}))



  

p <- ggplot(x[!x$karyotype=="others",],aes(x=time,y=frequency))+
  facet_grid(rows=vars(pm),cols=vars(rep))+
  geom_area(aes(fill=karyotype),show.legend = F)
plot(p)




```

Messing round with dimred

```{r}

optimf <- function(pars){
  y <- matrix(c(0,0,pars),ncol=2,byrow=T)
  d2 <- as.matrix(dist(y))
  sum((d[lower.tri(d)] - d2[lower.tri(d2)])^2)
}

x <- rbind(c(0,0,0),
           c(1,0,0),
           c(0.5,0.))
d <- as.matrix(dist(x))

pars <- runif(2*nrow(d)-2)


```
```{r}
#set.seed(42)
stepr <- function(ki){
  ni <- gen_all_neighbours(list(ki),as.strings = F)
  #ni <- ni[apply(ni,1,function(nij) !0%in%nij),]
  deltaf <- c(predict(m,ni))-c(predict(m,matrix(ki,nrow=1)))
  if(max(deltaf)<0) return(ki)
  return(ni[which.max(deltaf),])
}
optimf <- function(pars){
  y <- matrix(c(0,0,pars),ncol=2,byrow=T)
  d2 <- as.matrix(dist(y))
  sum((d[lower.tri(d)] - d2[lower.tri(d2)])^2/d[lower.tri(d)]^2)
}
optimf2 <- function(pars,d1){
  d2 <- apply(y,1,function(yi) sqrt(sum((pars-yi)^2)))
  sum((d1-d2)^2/d1^2)
}

m <- readRDS("salehi_data/03_inference_output/hTERTa/krig.Rds")
x <- readRDS("salehi_data/03_inference_output/hTERTa/frequent_clones.Rds")
kx <- do.call(rbind,lapply(rownames(x), function(xi) as.numeric(unlist(strsplit(xi,split="[.]")))))
## find all distance 1 neighbours of the input data:
n1 <- gen_all_neighbours(rownames(x))
## find all distance 2 neighbours:
n2 <- gen_all_neighbours(apply(n1,1,paste,collapse="."))
k <- rbind(kx,n1,n2)
k <- k[apply(k,1,function(ki) prod(ki>0)==1),]

peaks <- do.call(rbind,lapply(1:nrow(kx), function(i){
  ki <- k[i,]
  kj <- rep(0,length(ki))
  while(!prod(ki==kj)) {
    kj <- ki
    ki <- stepr(ki)
    #print(kj)
  }
  return(ki)
}))
peaks <- unique(peaks)
d <- as.matrix(dist(peaks))
opt <- lapply(1:200, function(i){
  pars <- 10*runif(2*nrow(d)-2)
  opt <- optim(pars,optimf)
})

v <- sapply(opt,function(o) o$value)
opt <- opt[v==min(v)]
y <- matrix(c(0,0,opt[[1]]$par),ncol=2,byrow=T)
df <- data.frame(y)
df$k <- apply(peaks,1,paste,collapse="")

Z <- k[sample(1:nrow(k),1000),]

df2 <- do.call(rbind,lapply(1:nrow(Z), function(i){
  #print(i)
  z=Z[i,]
  if(paste(z,collapse="")%in%apply(peaks,1,paste,collapse="")) return(NULL)
  d <- as.matrix(dist(rbind(z,peaks)))
  d1 <- d[1,-1]
  d <- d[-1,-1]
  pars <- y[which(d1==min(d1))[1],] #initial guess
  pars <- optim(pars,optimf2,d1=d1)$par
  data.frame(X1=pars[1],X2=pars[2],k=paste(z,collapse=""))
}))





df3 <- rbind(df,df2)
tmp <- do.call(rbind,lapply(df3$k, function(ki) as.numeric(unlist(strsplit(ki,split="")))))
tmp <- predict(m,tmp)
df3$fitness <- tmp

tmp <- akima::interp(df3$X1,df3$X2,df3$fitness,duplicate="mean",nx=100,ny=100)

lvis <- tmp[[3]]
rownames(lvis)<-tmp$x
colnames(lvis)=tmp$y
tmp <- reshape2::melt(lvis)
colnames(tmp) <- c("X1","X2","fitness")
p <- ggplot(df,aes(x=X1,y=X2))+
  geom_point()+
  ggrepel::geom_text_repel(aes(label=k))
p
p <- ggplot(tmp,aes(x=X1,y=X2))+
  geom_raster(aes(fill=fitness))+
  geom_point(data=df)+
  ggrepel::geom_text_repel(data=df,aes(label=k))+
  scale_fill_viridis_c()
p
```

