---
title: "README"
author: "Richard J Beck"
date: "`r Sys.Date()`"
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/projects/008_birthrateLandscape/ALFA-K/")
```

Data in this folder was derived from longitudinal sequencing data obtained from *Salehi, Sohrab, et al. "Clonal fitness inferred from time-series modelling of single-cell cancer genomes." Nature 595.7868 (2021): 585-590.* In particular, Extended Data Fig. 1 in their paper is a useful reference for relating the data in this folder to the underlying experimental procedures.

The data objects provided here contain karyotype frequency data formatted as follows:


```{r}

x <- readRDS("example_data/hTERTb.Rds")
head(x$x)

```
Note that the column names refer to the passage number as indicated in Salehi et. al. The data we received contained copy number calls for a multitude of genomic segments per chromosome. We reduced these data to whole-chromosome resolution by assigning each chromosome the modal copy number of all of it's segments. 

For more information about the formatting of these data, see example_1
 
