---
title: "Example 2"
author: "Richard J Beck"
date: "2023-10-13"
output:
  md_document:
    variant: markdown_github
---

**This example illustrates a leave-one-out cross validation procedure to validate ALFA-K if no ground truth data is available.**

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/projects/008_birthrateLandscape/ALFA-K/")
```

```{r,warning=FALSE,message=FALSE}
source("utils/ALFA-K.R")

```
Below are copy number calls for hTERTa data set already formatted for alfa-k. Note the units of "time" are passage numbers (we do not know how long the passages were). We assumed passage length of 5 days, hence dt is set to 5 so fitness will be in units of day^(-1).
```{r}
x <- readRDS("example_data/hTERTa.Rds")
head(x$x)
print(x$dt)
```
Run alfa-k on data
```{r,warning=FALSE,message=FALSE,result=FALSE}

opt <- alfak(x,min_obs = 25)

```
Validation
```{r,warning=FALSE,message=FALSE,result=FALSE}

## recover frequent clones
xfq <- opt$xo[opt$xo$id=="fq",]
##leave one out cross validation - use ALFA-K to predict fitness of frequent clones left out of the training data.
df <- do.call(rbind,lapply(1:nrow(xfq),function(i) optim_loo(i,x,xfq)))

##positive correlation indicates predictive value
plot(df$pred,xfq$f_est)
cor(df$pred,xfq$f_est)
```

