---
title: "Example 1"
author: "Richard J Beck"
date: "2023-10-10"
output:
  md_document:
    variant: markdown_github
---

**This example illustrates how to fit a landscape using ALFA-K and run an ABM simulation using the fitted landscape**

We will fit ALFA-K to data generated using the ABM simulated with a random artificial fitness landscape. First, ensure ABM is compiled, root dir is set to correct location of ALFA-K, and the cpp_source is correctly set to the path of the ABM executable file.
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/projects/008_birthrateLandscape/ALFA-K/")
```

```{r}
cpp_source <- "ABM/bin/ABM"
```

```{r,warning=FALSE,message=FALSE}
source("utils/sim_setup_functions.R")
source("utils/ALFA-K.R")
dir.create("examples/example_1/sim_output")
```
Setup ABM simulation with random fitness landscape

The gen_replicate() function can be used to setup the directory structure required to run an ABM simulation. 

**Arguments**

**Nchrom:** number of chromosomes for each cell in simulation

**wavelength:** controls complexity of GRF fitness landscape. Lower wavelengths are more complex. 

**sweep_dir:** output directory for sim

**cpp_source:** path to ABM executable

**p:** missegregation rate

**Value**

**cpp_run_cmd:** command necessesary to run ABM simulation

**sweep_id:** name of output folder that has been generated

The following chunk sets up and runs ABM simulation:
```{r, results=FALSE}
Nchrom <- 22
wavelength <- 0.8
misrate <- 0.00005
sweep_dir <- "examples/example_1/sim_output/"

info <-  gen_replicate(Nchrom=Nchrom,wavelength = wavelength,p = misrate,sweep_dir = sweep_dir,cpp_source = cpp_source)
system(info$cpp_run_cmd,show.output.on.console = F)
```
The ABM simulation directory folder contains several files and folders. The config file contains a lot of the information about the ABM sim. The output_dir entry tells us where the sim output will be saved. 

```{r}
list.files(paste0(sweep_dir,info$sweep_id))
config <-readLines(paste0(sweep_dir,info$sweep_id,"/config.txt"))
print(config)
output_dir <- unlist(strsplit(config[grepl("output_dir",config)],split=","))[2]

```

The output directory contains numbered subfolders corresponding to repeated ABM runs with identical parameter sets. If we run the ABM simulation again a new subfolder is created. 

```{r}

list.files(output_dir)
system(info$cpp_run_cmd,show.output.on.console = F)
list.files(output_dir)

```

Each subfolder contains numbered csv files containing population samples at indicated timepoints. There is also a summary file which contains population size of the ABM at all timepoints simulated.

```{r}
output_subdir <- paste(output_dir,list.files(output_dir)[1],sep="/")
head(list.files(output_subdir))
pop.size <- readLines(paste0(output_subdir,"/summary.txt"))
plot(pop.size)
```

Process output from ABM simulation:

The proc_sim() function compiles output from the ABM simulation into a format amenable to further analysis

**Arguments**

**dir:** directory containing csv output files from ABM simulation

**times:** numeric vector containing times to sample output from ABM simulation. For each user-supplied time supplied, the closest timepoint among ABM csv output files is selected. 

**Value**

**x:** matrix with rows corresponding to karyotypes and columns corresponding to timepoints. Values indicate the number of sampled cells. 

**pop.fitness:** growth rate of the sampled ABM population at each sampled timepoint 

**dt:** conversion factor between output timepoints and time in days. E.g. if dt=0.1, timepoint 400 = 40 days. 

**clone.fitness:** fitness of each karyotype in matrix x. values are ordered according to rows of x. 




```{r,message=FALSE, warning=FALSE}
x <- proc_sim("data/htert/train/00000/",times = seq(0,2800,400))
x <- proc_sim(dir=output_subdir,times = seq(0,2800,400))
print(head(x$x[order(rowSums(x$x),decreasing = T),]))
```

Estimate fitness landscape:

```{r,message=FALSE, warning=FALSE}

fit <- alfak(x,min_obs = 20)
xmat <- do.call(rbind,lapply(rownames(fit$xo), function(i){
    as.numeric(unlist(strsplit(i,split="[.]")))
  }))
f_est <- predict(fit$fit,xmat)
```

Convert Krig to format readable by ABM and save:
```{r}

knots <- fit$fit$knots
cc <- fit$fit$c
d <- fit$fit$d
fscape <- rbind(cbind(knots,cc),c(d))
fitted_landscape_path <- paste0(sweep_dir,info$sweep_id,"/fitted_landscape.txt")
write.table(fscape, fitted_landscape_path,row.names = FALSE,col.names=FALSE,sep=",")

```

Generate a new config file to use fitted landscape. Run ABM.
```{r, results=FALSE}

base_config <- readLines(paste0(sweep_dir,info$sweep_id,"/config.txt"))
new_config <- modify_config("fitness_landscape_type","krig",base_config)
new_config <- modify_config("fitness_landscape_file",fitted_landscape_path,new_config)

# we will save output from the fitted landscape in the test subfolder
test_dir <- paste0(sweep_dir,info$sweep_id,"/test") 
new_config <- modify_config("output_dir",test_dir,new_config)

new_config_path <- paste0(sweep_dir,info$sweep_id,"/fitted_landscape_config.txt") 
writeLines(new_config,new_config_path)
system(paste(cpp_source,new_config_path),show.output.on.console = F)
```

Run ABM with custom karyotype composition

```{r, results=FALSE}
custom_pop_output <- paste0(sweep_dir,info$sweep_id,"/custom_pop")

dir.create(custom_pop_output)
k1 <- rep(2,22)
k2 <- rep(4,22)
pop <- rbind(k1,k2)

n_init <- c(50000,50000)
pop <- cbind(pop,n_init)

custom_pop_path <- paste0(sweep_dir,info$sweep_id,"/custom_pop.txt")
write.table(pop,custom_pop_path,sep=",",row.names = F,col.names = F)

config <- readLines(paste0(sweep_dir,info$sweep_id,"/config.txt"))
config <- c(config,paste0("population_file,",custom_pop_path))
config <- modify_config("output_dir",custom_pop_output,config)
custom_config_path <- paste0(sweep_dir,info$sweep_id,"/custom_pop_config.txt")
writeLines(config,custom_config_path)
system(paste(cpp_source,custom_config_path),show.output.on.console = F)
```

